---
title: "Homework 2"
author: "Eileen Shea"
date: "October 5, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Problem 1

First we need to read and clean the NYC transit data, while also retaining only the necessary variables and converting the _entry_ variable from character to logical:

```{r}
subway_data = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:route11, entry, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

The dataset we have right now is a subset of NYC Transit data that relates information about entrances and exits for each subway station in NYC. Variables that have been kept as part of our dataset are: line, station_name, station_latitude, station_longitude, route1-11, entry, vending, entrance_type, and ada. These data encompass numerous data types, including character variables (e.g. line, station_name), double variables (e.g. station_latitude, station_longitude), integer variables (e.g. route8, route11), and logical variables (e.g. ada, entry). So far, the data cleaning steps have involved turning all of the variable names into lower snake case, getting rid of extraneous variables, and converting between character and logical variable for _entry_. The dataset consits of 1868 rows and 19 columns. Right now, the dataset is not particularly tidy, especially due to the fact that some columns are values instead of variable names (i.e. route1-11 columns). 




