---
title: "Homework 2"
author: "Eileen Shea"
date: "October 5, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Problem 1

First we need to read and clean the NYC transit data, while also retaining only the necessary variables and converting the _entry_ variable from character to logical:

```{r}
subway_data = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:route11, entry, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

The dataset we have right now is a subset of NYC Transit data that relates information about entrances and exits for each subway station in NYC. Variables that have been kept as part of our dataset are: line, station_name, station_latitude, station_longitude, route1-11, entry, vending, entrance_type, and ada. These data encompass numerous data types, including character variables (e.g. line, station_name), double variables (e.g. station_latitude, station_longitude), integer variables (e.g. route8, route11), and logical variables (e.g. ada, entry). So far, the data cleaning steps have involved turning all of the variable names into lower snake case, getting rid of extraneous variables, and converting between character and logical variable for _entry_. The dataset consits of 1868 rows and 19 columns. Right now, the dataset is not particularly tidy, especially due to the fact that some columns are values instead of variable names (i.e. route1-11 columns). 

First, we are curious about how many distinct stations there are:

```{r}
distinct_sub = subway_data %>% 
  distinct(line, station_name, .keep_all = TRUE)
```

From the above code chunk, we learn that there are **465** distinct subway stations.

Next, we are curious about how many stations are ADA compliant:

```{r}
sum(distinct_sub$ada)
```

It turns out that there are only **84** ADA compliant subway stations out of that bunch.

We are then curious about what proportion of station entrances / exits without vending allow entrance:

```{r}
no_vending = subway_data %>% 
  filter(vending == "NO")

sum(no_vending$entry)/length(no_vending$entry)
```

From this code chunk we learn that the proportion of station entrances / exits without vending that allow entrance is **0.377**.

Now it is important to reformat the dataset so that route number and route name are distinct variables:

```{r}
subway_tidy_data = gather(subway_data, key = rt_number, value = route_name, route1:route11) %>% 
  separate(rt_number, into = c("remove", "route_number"), sep = 5) %>% 
  select(everything(), -remove)
```

We also are interested in identifying how many distinct stations serve the A train:

```{r}
A_train = subway_tidy_data %>% 
  distinct(line, station_name, .keep_all = TRUE) %>% 
  filter(route_name == "A")

nrow(A_train)
```

The A train according to this dataset serves **60** distinct stations.

Of the stations that serve the A train, we are interested in which are ADA compliant:

```{r}
sum(A_train$ada)
```

Out of 60 A train stations, **17** are ADA compliant.






